---
title: "Lab Exercises for Week 4"
author: ""
date: "January 30, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Lab This Week

1. Load the tidyverse library which contains a dataset called "diamonds" which is a dataset containing the prices and other attributes of diamonds.
2. Give two commands that you could use to get a feel for what the diamonds dataset contains (e.g., the variables it contains, the values the variables can take)
3. The following barplot was created using the diamonds data. Provide the code to recreate the plot.

```{r echo = FALSE, include=TRUE}
ggplot(data = diamonds) +
  geom_bar(
    mapping = aes(x=cut, fill = clarity),
    position = "dodge"
  )


```


4. What command would you use to save the subset of the diamonds data for which the cut is "Ideal" and the clarity is "SI1" or (inclusive) "SI2" or "" to a new variable which we'll name "df"? Can you think of another way to do the same thing? (Hint: use "%in%")

5. How would you alter your code in question 4 to use the same filter (i.e., we're still interested in the subset of data for which the cut is "Ideal" and the clarity is "SI1" or (inclusive) "SI2") but to only select the carat column instead of saving all the columns?