---
title: "Week 5 Exercises"
author: ""
date: "February 13, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab This Week

Your job is to master the gollowing tidyverse functions: group_by, arrange, mutate, select, filter, and summarise.

We will be working with a dataset in the "DALEX2" pacage. Let's install the package.

```{r echo = TRUE, message = F, warning = F, eval = F}
# Install package
install.packages('DALEX2')
```

Now let's load both the DALEX2 package and the tidyverse

```{r echo = TRUE, message = F, warning = F}
library(DALEX2)
library(tidyverse)
```

Finally, let's load our dataset, we will be working with a datset on dragons. Run the following code to load the datset.

```{r echo = TRUE, message = F, warning = F}
data(dragons)
View(dragons)
```

Now let's explorre this dataset a bit. Suppose we're interested in computing the number of dragons of each color. We can do that with the following code.

```{r echo = TRUE, message = F, warning = F}
# Count dragons of each color
dragon_count <- dragons %>%
  group_by(colour) %>%
  summarise(Number = n()) %>%
  as_tibble()
dragon_count
```

Now, suppose we want to look at only the pretty dragons (less than 5 scars and 5 missing teeth). We can filter the dragons with the following.

```{r echo = TRUE, message = F, warning = F}
# Dragons less than 5 scars
pretty_dragons <- dragons %>%
  filter(scars < 5 & number_of_lost_teeth < 5) %>%
  as_tibble()
pretty_dragons
```

Your job is to perform the following tasks:

1. Create a table of the 5 dragons with the most scars, displaying their year of birth, scars, color, and number of lost teeth. (Hint: Use function top_n)

2. Among the dragons that have greater than 20 scars, what number are blue?

3. On **average**, dragons of which color lost the most teeth. Create a table displaying these averages.

Once you have mastered the dragons dataset, you may move on to your lab. Good luck.