---
title: "Week 13 Exercises"
author: ""
date: "April 17, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab This Week

You have to individually:

  * write a function with another function as an argument
  * use a new data manipulation function called "map"  
  
Note that you are essentially doing the same thing in both of these tasks, however we are making you implement both methods so that you can appreciate the map function more.

## Preliminary
  
Let's start with loading our packages and generate a toy dataset to work with.

```{r echo = T, message = F, warning = F}

# Load tidyverse
library(tidyverse)

# Create data frame
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
```

## Functions within functions

We will be working with the example from section 21.4 in *R for Data Science*. Suppose we want to create a function that computes column summary statistics, but we want the statistic to be a parameter of the function. We can do this by writing a function, with another function as an argument. Let's see how we can do this.

```{r echo = T, message = F, warning = F}

# Function to compute column summary
col_summary <- function(df, fun) 
{
  ## Inputs:
  # df: A data frame object
  # fun: Function for statistic, ex. mean, median, sd, etc.
  
  ## Ouputs:
  # out: Vector of summary stats for each column
  
  # Empty vector to store values - same length as ncol(df)
  out <- vector("double", length(df))
  
  # Loop through columns of data frame
  for (i in seq_along(df)) {
    
    # Compute summary statistc of column
    out[i] <- fun(df[[i]])
  }
  
  # Return summary statistics
  return(out)
}
```

Now let's call this function on out tibble that we generated to compute the mean.

```{r echo = T, message = F, warning = F}

# Compute mean column summary
col_summary(df, mean)
```

**Exercise: Change the input function to compute a different summary statistic.**

## map Functions

For alot of things that usually require for loops, we can use map functions from the purrr package to achieve the same result. This allows us to avoid writing the for loop like in the last example. There are five main map functions, one for each specified output type. They are:

* map()  
* map_lgl()  
* map_int()  
* map_dbl()  
* map_chr()

Let's check out the folowing example, taken from *R for Data Science* section 21.5.

```{r echo = T, message = F, warning = F}

# Use a map function compute mean of each column
df %>%
  map_dbl(mean) 

```

**Exercise: Adapt the map function to compute the median and standard deviation of each column.**

These are just basic examples. Let's get more understanding of the map function. Notice the syntax is different in the next example, however we achieve the same result.

```{r echo = T, message = F, warning = F}

# Use map function to compute mean
values <- df %>%
  map(function(df) mean(df))
values
```

Notice that the output of the general map function is a list. We can index the list in the folowing way.

```{r echo = T, message = F, warning = F}
# Index by name
values$a

# Index by position
values[[1]]
```

Or we could just use the map_dbl function instead to avoid the list altogether.

```{r echo = T, message = F, warning = F}
# Use map_dbl
values <- df %>%
  map_dbl(function(df) mean(df))
values
```

Finally, you can also define map functions to work with your own defined functions.

```{r echo = T, message = F, warning = F}

# Function to compute add 10 mean
add10mean <- function(x)
{
  x <- x + 10
  return(mean(x))
}

# Way 1
df %>%
  map_dbl(add10mean)

# Way 2
df %>%
  map_dbl(function(df) add10mean(df))
```

**In your lab this week, you need to show that you have ability in doing the above two tasks. Your use of maps and functions within functions don't necessarily have to be related to the permutation test part of the lab.**